Exploitation Techniques and Mitigations

### 2016-01-16 docs/paper.tex
Paper: Draft format string exploits

\subsection{Endianness}

@@

\subsection{Calling Convention}

@@

>>>>

\subsection{Endianness}
\label{sub:endianness}

@@

\subsection{Calling Convention}
\label{sub:calling_convention}

@@

The first exploitation technique we will discuss builds upon the interpretation
of format strings. \texttt{printf} is a C function of the standard library
which will interpret such strings and print them to \texttt{stdout}. As the
name already tells you, the supplied string contains \textit{formatter}
describing how to actually handle additional arguments. If you are unfamiliar
with \texttt{printf} please have a look at the man
page\footnote{\url{http://linux.die.net/man/3/printf}}.

Taking a closer look at \texttt{printf} we can see that its first argument is a
format string followed by a variable number of additional arguments. In C you
don't know how many arguments have been supplied when a function with a
variable number of arguments is called. Some instances work around this by
taking an argument count as their first argument, others expect you to
terminate with a special symbol (usually \texttt{NULL}). \texttt{printf} uses
the format string to derive how many arguments have been supplied. Calling
\texttt{printf}, for example, with the string \texttt{"\%d + \%d = \%d"}
assumes that (at least) three arguments have been provided.

\begin{listing}[h!]
    \begin{minipage}[t]{0.47\textwidth}
        \cfile{../format_string/main.c}
    \end{minipage}
    \begin{minipage}[t]{0.47\textwidth}
        \begin{pre*}{autogobble}
            > echo foobar | ./main
            You entered:
            foobar
        \end{pre*}
        \medskip
        \begin{pre*}{autogobble}
            > echo AAAABBBB | ./main
            correct
        \end{pre*}
        \medskip
        \begin{pre*} {autogobble}
            > echo '%08x' | ./main
            You entered:
            bfd98ed4
        \end{pre*}
    \end{minipage}
    \caption{Program vulnerable to Format String Exploits}
    \label{src:format_string_exploit}
\end{listing}

The exploit comes from the notion that a format string provided by an attacker
gets interpreted. The program shown in \cref{src:format_string_exploit} will
take an arbitrary string from \texttt{stdin} and pass it on to \texttt{printf}.
For simple inputs (not containing formatter) this works fine. But as soon as
formatter are provided, \texttt{printf} is going to access the locations where
the corresponding arguments \emph{would} be located. From the calling
convention described in \cref{sub:calling_convention} we know that these
arguments would be located on the stack, therefore \texttt{printf} will print
whatever lies on the stack.

To fully exploit the provided example, note that an attacker in this scenario
wants to get a hold of the hardcoded password stored in \texttt{passwd}. Since
local variables are placed on the stack \texttt{printf} will be able to read
the password if enough formatters are provided:

\begin{pre*} {autogobble}
    > python -c 'print "%08x." * 10' | ./main
    bf920c14.00000064.b77de29e.00000000.00000000.b77fedf8.bf920d94.00000000.41414141.42424242.
\end{pre*}

Here we use Python to craft the format string for us. As we can see the
password is printed (ASCII encoded). Byte order is swapped because of
endianness (see \cref{sub:endianness}). Apart from the password we also gather
a bunch of pointers, these can be used later on to break ASLR (see
\cref{sub:info_leak}).

We would like to point the reader to the book \textit{Hacking: The Art of
Exploitation} \cite[pp.~167]{art_of_exploitation} for more details about this
technique. We will come back to this technique later on to show that
\texttt{printf} enables even more sophisticated attacks.
